---
title: "p8105_HW2_rat2134"
author: "Robert Tumasian"
date: "9/28/2019"
output: github_document
---

```{r, message=FALSE}
#Load required packages
library(readxl)
library(janitor)
library(tidyverse)
library(dplyr)
library(tibble)
library(knitr)
```

# Problem 1
```{r}
#Import Mr. Trash Wheel dataset 
Mr_Trash_Data = 
  read_excel("./HW2_datasets/Mr_Trash_Wheel.xlsx",
             range = cell_cols("A:N")) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  mutate(sports_balls = as.integer(
    round(sports_balls, 
          digits = 0)))
```

```{r}
#Import and clean precipitation data for 2017 and 2018
Precipitation_data_2017 =
  read_excel("./HW2_datasets/Mr_Trash_Wheel.xlsx",
             sheet = 6,
             cell_rows(3:14),
             col_names = c("Month", "Total_Precipitation")) %>%
  mutate(Year = 2017)

Precipitation_data_2018 =
  read_excel("./HW2_datasets/Mr_Trash_Wheel.xlsx",
             sheet = 5,
             cell_rows(3:14),
             col_names = c("Month", "Total_Precipitation")) %>%
  mutate(Year = 2018)

#Combine data sets
Precipitation_data_2017_and_2018 = 
  bind_rows(Precipitation_data_2017, Precipitation_data_2018) %>%
  mutate(Month = month.name[Month]) %>%
  select(Year, Month, Total_Precipitation, everything())
```

`Mr_Trash_Data` has `r dim(Mr_Trash_Data)[1]` observations and `r dim(Mr_Trash_Data)[2]` variables:

* Dumpster number
* Month
* Year
* Date of dumpster check (YYYY-MM-DD)
* Weight (in tons) and volume (in cubic yards) of trash in dumpster
* Quantity of plastic bottles, polystyrene, cigarette butts, glass bottles, grocery bags, chip bags, and sports balls found in the dumpster
* Number of homes powered (each ton of trash equates to on average 500 kilowatts of electricity, and an average household will use 30 kilowatts per day)

`Precipitation_data_2017_and_2018` has `r dim(Precipitation_data_2017_and_2018)[1]` observations and `r dim(Precipitation_data_2017_and_2018)[2]` variables:

* Year
* Month
* Total Precipitation (in inches)

With the data available, the total precipitation in 2018 was `r sum(pull(filter(Precipitation_data_2017_and_2018, Year == 2018), Total_Precipitation))` inches. In addition, the median number of sports balls found in the dumpsters in 2017 was `r median(pull(filter(Mr_Trash_Data, year == 2017), sports_balls))`.

# Problem 2
```{r, message=FALSE}
#Import and clean pols-month dataset
Pols_month = 
  read_csv("./HW2_datasets/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.factor(month)]) %>%
  #In addition to 0s and 1s in the prez_gop variable, there are also 2s
  #All 2s will be reassigned to 1s
  mutate(prez_gop = ifelse(prez_gop == 1 | prez_gop == 2, 1, 0)) %>%
  mutate(president = recode(prez_gop, `0` = "dem", `1` = "gop")) %>%
  select(-c(prez_dem, prez_gop, day))
```

```{r, message=FALSE}
#Import and clean snp dataset
snp =
  read_csv("./HW2_datasets/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("month", "day", "year")) %>%
  arrange(year, month) %>%
  select(-day) %>%
  mutate(month = month.name[as.factor(month)]) %>%
  select(year, month, everything())
```

```{r, message=FALSE}
#Import and clean unemployment dataset
unemployment_data = 
  read_csv("./HW2_datasets/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_percentage"
  ) %>%
  rename(year = Year) %>%
  arrange(year, month) %>%
  mutate(month = month.name[as.factor(month)]) %>%
  mutate(year = str_c(year))
```

```{r}
#Merging all three datasets
snp_and_pols = left_join(Pols_month, snp, by = c("year", "month"))

snp_and_pols_and_unemployment = left_join(snp_and_pols, 
                                          unemployment_data, 
                                          by = c("year", "month"))
```

The `Pols_month` dataset contains `r dim(Pols_month)[1]` observations and `r dim(Pols_month)[2]` variables:

* Year
* Month
* Number of republican governors, senators, and representatives
* Number of democratic governors, senators, and representatives
* The political party of the serving President

This data is available from January 1947 to June 2015.

The `snp` dataset contains `r dim(snp)[1]` observations and `r dim(snp)[2]` variables:

* Year
* Month
* The closing value of the S&P stock index on the associated date

This data is available from January 1950 to October 2015.

The `unemployment_data` dataset contains `r dim(unemployment_data)[1]` observations and `r dim(unemployment_data)[2]` variables:

* Year
* Month
* Percentage of unemployment

This data is available from January 1948 to September 2015.

Lastly, the `snp_and_pols_and_unemployment` dataset (merged by year and month) contains `r dim(snp_and_pols_and_unemployment)[1]` observations and `r dim(snp_and_pols_and_unemployment)[2]` variables. All the variables in this combined dataset contain those in the `Pols_month`, `snp`, and `unemployment_data` datasets. This data is available in different time intervals for different variables, which explains why there are some `NA` values.

# Problem 3
```{r, message=FALSE}
#Import and clean baby names data
popular_baby_names = 
  read_csv("./HW2_datasets/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = replace(ethnicity, 
                             ethnicity == "ASIAN AND PACI", 
                             "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = replace(ethnicity, 
                             ethnicity == "WHITE NON HISP", 
                             "WHITE NON HISPANIC"),
         ethnicity = replace(ethnicity, 
                             ethnicity == "BLACK NON HISP", 
                             "BLACK NON HISPANIC")) %>%
  mutate(ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name),
         gender = str_to_lower(gender)) %>%
  distinct()
```

```{r}
#Table to show rank in popularity of "Olivia" from 2011 to 2016
Olivia_rank_table = 
  filter(popular_baby_names, childs_first_name == "olivia") %>%
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank) %>%
  rename("Ethnicity" = ethnicity)

kable(Olivia_rank_table)
```

```{r}
#Table showing most popular male baby name over time
Most_popular_male_baby_name = 
  filter(popular_baby_names, gender == "male" & rank == 1) %>%
  select(-c(gender, count, rank)) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name) %>%
  rename("Ethnicity" = ethnicity)

kable(Most_popular_male_baby_name)
```

```{r}
#Obtain data for all male, white non-hispanic baby names in 2016
Male_White_Non_Hispanic_Names_2016 = 
  filter(popular_baby_names, gender == "male" & 
           ethnicity == "white non hispanic" &
           year_of_birth == 2016)

#Create scatterplot of number of children vs. rank
children_vs_rank_plot = 
  ggplot(Male_White_Non_Hispanic_Names_2016, 
         aes(x = rank, y = count)) +
  geom_point() +
  xlab("Rank") +
  ylab("Number of Children") +
  ggtitle("Number of Children with a Name vs. Rank")

plot(children_vs_rank_plot)
```